Designing Qusecs Optimally: Quasi-uniform or Self-similar Layered Materials as Recursively Decomposed Solutions of Geometric Constraint Systems

### 2014-12-18 doc/GMP2015/s04_recombination.tex
Attempt to unify text decoration/emphasis. Make sure to pull, affects all tex files.

In this section, we consider the {\em optimal recombination}
problem of combining specific solutions of
subsystems in a DR-plan into a solution of their parent system $I$, (without
loss of generality, at the top level of the DR-plan). In the case of isostatic
qusecs, the parent system $I$ is isostatic (the root of the DR-plan), and we
seek  {\em solution(s) (among a finite large number of solutions)
with a specific orientation or chirality}.
In the case of underconstrained qusecs the subsystems are the multiple roots
of the DR-plan, the parent system $I$ is underconstrained, and we typically seek
an efficient algorithmic description of {\em connected component(s) of solutions with a specific orientation or
chirality}.

The main barrier in recombination when given an optimal DR-plan (of smallest possible
size or max fan-in),  is that the number of children (resp. number of
roots) - and correspondingly the  size and complexity of the (indecomposable) algebraic
system $I$ to be solved -
could be arbitrarily large as a function of the size of the input system.
This difficulty can persist even after optimal parametrization of the
indecomposable system $I$ as in \cite{XX}
to minimize its algebraic complexity.

In the following, we formulate the problem of {\em optimal
modification},
of the indecomposable
algebraic system $I$ into a decomposable system with a small DR-plan (low
algebraic complexity).
Leveraging recent results on {\em Cayley
configuration spaces}, our approach to the optimal modification problem
achieves the following: {\em (a) Small DR-plan.} We obtain a  parameterized family of systems
$I_{\lambda_F}$ -  one for each value $\lambda_F$ for the parameters $F$,  all of which have
small DR-plans. Thus, given a value $v$ for $\lambda_F$, the system $I_v$ can
potentially be solved or realized easily once the orientation type of the solution
is known  (when the DR-plan size is small enough).
{\em (b) Solution preservation.} Moreover, the union of solution spaces of the systems in the family
$I_{\lambda_F}$
is guaranteed to contain
all of $I$'s solutions. {\em (c) Efficient search.} Finally, this union can be projected into the
so-called {\em Cayley} or distance parameter space $\lambda_F$ that is convex
or otherwise easy to traverse in order to search for $I$'s solution
(or connected component)
of the desired orientation type.
For the case when the modification is bounded, this approach provides an efficient
algorithm for recombination. We first define the decision version of the
problem of optimal modification
for decomposition. The standard optimization versions are straightforward.

\medskip\noindent{\bf Optimal Modification for Decomposition (OMD) Problem.} Given a generically
independent graph $G = (V,E)$
with no non-trivial proper isostatic subgraph (indecomposable), and 2 constants $k$ and $s$,
does there exist a set of at most $k$  edges $E_1$ and a set of nonedges $F$
such that the modified graph $H = (V, E\setminus E_1 \cup F)$ has a DR-plan
of size at most $s$?  The OMD$_k$ problem is  OMD where $k$ is a fixed bound (not
part of the input).   We say that such a tuple $(G,s)$ is a member of the
set OMD$_k$. We loosely refer to graphs $G$ as OMD with appropriately small $k,s$ or OMD$_k$ graphs
with appropriately small $s$.

It is immediately clear that indecomposable graphs $G$ that belong in OMD$_k$ for small $k$ and
$s$  lend themselves to modification  into decomposable graphs
satisfying Criteria (a) and (b) above.
However, it is not clear how Criterion (c) is met by OMD graphs.
%
\subsection{Using Convex Cayley Configuration Spaces}
%
\ref{2-tree-reduction}
Next we provide the necessary background to describe a specific approach for achieving the requirements
(a)-(c) mentioned above, by restricting the class of reduced graphs $G' =
G\setminus E_1$ and their isostatic completions $H$ in the above definition of the
OMD problem, and using a key theorem of Convex Cayley configuration
spaces \cite{XX}. This theorem characterizes the class of graphs $H$ and non-edges $F$ (Cayley
parameters), such that the set of vectors $\lambda_F$ of  attainable lengths of
the nonedges $F$
is always convex for any
given lengths $\delta$ for the edges of $H$ (i.e, over all the realization
s of the bar-joint constraint system or linkage $(H,\delta)$ in 2 dimensions).
This set is called the (2-dimensional) {\em Cayley configuration space}
of the linkage $(H,\delta)$ on the Cayley parameters $F$, denoted
$\Phi_F(H,\delta)$ and can be viewed as a ``projection'' of the cartesian
realization space of $(H,\delta)$ on the Cayley parameters $F$.
Such graphs $H$ are said to have {\em convexifiable Cayley configuration spaces
with parameters $F$} (short: {\em convexifiable}). To state the theorem, we first have to define the
notion of {\em 2-sums} and {\em 2-trees}.
Let $H_1$ and $H_2$ be two graphs on disjoint sets of vertices $V_1$ and
$V_2$, with edge sets $E_1$ and $E_2$ containing edges $(u,v)$ and $(w,x)$
respectively.  A {\em 2-sum} of
$H_1$ and $H_2$ is a
graph $H$  obtained by taking the union of $H_1$ and $H_2$ and identifying $u=w$ and $v=w$.
In this case, $H_1$ and $H_2$ are called {\em 2-sum components} of $H$.
A {\em minimal 2-sum component} of $H$ is  one that cannot be further split
into 2-sum components.
A {\em 2-tree} is recursively obtained by taking
a 2-sum of two 2-trees, with the base case of 2-tree being a triangle.
A {\em partial 2-tree} is a 2-tree minus some edges.
Partial 2-trees are characterized as having $K_4$ as forbidden minor are also called series parallel graphs.

\begin{theorem}
    \label{convexcayley}
    \cite{XX}
    $H$ has a convexifiable Cayley configuration space  with parameters $F$
    if and only if for each $f\in F$  all the minimal 2-sum components
    of $H\cup F$ that contain both endpoints of $f$ are partial 2-trees.
    The Cayley configuration space $\Phi_F(H,\delta)$
    of a bar-joint system or linkage $(H,\delta)$ is a convex polytope.
    When $H\cup F$ is a 2-tree, the
    bounding hyperplanes of this polytope are triangle inequalities
    relating the lengths of edges of the triangles in $H\cup F$.
\end{theorem}

@@

\begin{theorem}
    \label{omdk}
Given an indecomposable graph $G$,
let $G'$ be a spanning partial 2-tree subgraph in $G$ with
$k$ fewer edges than $G$.
Then  $(G,2)$ belongs in the set OMD$_k$.
\end{theorem}

@@

We refer to such graphs $G$ in short as {\em $k$-approximately
convexifiable,} where the reduced graphs $G'$ and isostatic completions $H$
are convexifiable.
As observed earlier, since graphs such as $G$ are in OMD$_k$, Criteria (a) and (b)
are automatically met for small enough $k$.
Criterion $c$ is addressed
as described in the following efficient search procedure
which clarifies the dependence of the complexity on the number and ranges  of Cayley
parameters $F$.

\begin{theorem}
    \label{criterionc}
    [Efficient search]
    For an indecomposable, $k$-approximately convexifiable graph $G =
    (V,E)$,
   let $G' = (V,E' =E\setminus D)$ be a spanning partial 2-tree subgraph
    where $|D| \le  k$. Let  $F$ be a set of nonedges of $G$ such that
    $H = (V, E'\cup F)$ is a 2-tree.
    Each solution $p$ (or connected component of a solution space)
    of $(G,\delta)$
    of an orientation type $\sigma_p$ can be found in time $O(\log(W))$ where
    $W$ is the number of cells of desired accuracy (discrete volume) of
    the convex
    polytope $\Phi_F(G',\delta_E')$.  The (discrete) volume $W$ is exponential in
    $|F|$ and polynomial in the (discrete range) of the parameters in $F$.
\end{theorem}

\begin{proof}
    The cartesian realization
    space of $(H,<\delta_E', \lambda_F>)$ is
    computed easily with a DR-plan of size 2, and is the union of $2^t$
    solutions (modulo orientation preserving isometries) each with a distinct orientation type, where $t$ is the number of triangles in the 2-tree
    $H$;
    here $\delta_{S}$ is the restriction of the length vector $\delta$ to
    the edges in $S$.
    A desired solution $p$ (or connected component of a solution space)
    of $(G,\delta)$
    of an orientation type $\sigma_p$ can be found
    by a subdivided binary search   of the cartesian realization space of
    $(H, <\delta_{E'},\lambda_F>)$ of orientation type $\sigma_p$ -
    as $\lambda_F$ ranges over
    the discretized convex
    polytope $\Phi_F(G',\delta_E')$ with bounding hyperplanes described in
    Theorem \ref{convexcayley}.
    A solution $p$  is found  when the lengths for nonedges in $D$ match
    $\delta_D$.
\end{proof}

@@

\noindent{\bf Note:} A huge advantage of the convex Cayley method
is that it is completely unaffected when $\delta$ are intervals rather than exact values~\cite{GaoSitharam}.
%cite my paper with Heping!!!

\medskip\noindent{\bf Example}.
A graph $G=K_{3,3}$  cannot be decomposed into any nontrivial isostatic
graphs, i.e, its DR-plan has a root and 9
children corresponding to the 9 edges. Solving or recombining the system
$(G,\delta)$
corresponding to the root of
this DR-plan involves solving a quadratic system with 8 equations and variables.
Instead of simultaneously solving this system, we could instead use the fact
that $G=K_{3,3}$ is in OMD$_1$: remove the
edge $e$ in Figure \ref{fig:omd_k33_example} to give a partial 2-tree $G'$.
Now add the nonedge $f$ to give a 2-tree $H$ with a DR-plan of size 2.
The Cayley configuration space $\Phi_f(G', \delta_{E\setminus e})$
is a single interval of attainable length values $\lambda_F$ for
the edge $f$.
When $\delta$ is generic, i.e, does not admit collinearities or coincidences
in the realizations of $(G,\delta)$,
the realization space of $(H, <\delta_{E\setminus e}, \lambda_f>)$ has 16
solutions $q^p_{\lambda_f}$ (modulo orientation preserving isometries),
with distinct orientation types $\sigma_p$ 
(two orientation choices for each of the 4 triangles)
that can be obtained by solving a sequence of 4 single quadratics in 1
variable (DR-plan of size 2).
By subdivided
binary search
in the interval $\lambda_f \in
\Phi_f(G', \delta_{E\setminus e})$, the desired
solution $p$  of $(G,\delta)$ is found when the length of the nonedge
$e$ in  the realization
$q^p_{\lambda_f}$
is $\delta_e$.
%
\subsection{Optimized Modification by Enlarging the class of Reduced
Graphs}
\label{tdecomp}
It is possible in principle to decrease $k$ for a OMD$_k$ graph
(i.e, the number of edges to be removed
to ensure an isostatic completion that is decomposable with a small DR-plan) by
considering reduced graphs $G'$ (and modified graphs $H$) that come from a larger class than
partial 2-trees but nevertheless have convex Cayley configuration spaces at
least when the realization space is restricted to a sufficiently
comprehensive orientation type.
In particular, the so-called {\em tree-decomposable graphs of low Cayley
complexity} \cite{XX} include the partial 2-trees and many others that are
not partial 2-trees. These too result in DR-plans of size 2 or 3, putting $G$ in the class
OMD$_k$ and thus meeting Criteria (a) and (b). The Criterion (c) is met
- for example when $k=1$ -  because 1 dof Cayley configuration spaces of
linkages based on such graphs $G'$
are known to be single intervals
when a comprehensive orientation type $\sigma_p$ of the sought solution $p$ is given.
In addition, the bounds of these intervals are of low algebraic
complexity.
More precisely, the bounds  can themselves be computed using a DR-plan of size 2 or 3,
i.e, the computation of these bounds is tree-decomposable. Alternatively,
the bounds are in a simple quadratic or radically solvable extension field over the rationals,
or can be computed by solving a
triangularized system of quadratics.

>>>>

In this section, we consider the \dfn{optimal recombination} problem
of combining specific solutions of subsystems in a DR-plan into a
solution of their parent system $I$, (without loss of generality, at
the top level of the DR-plan). In the case of isostatic qusecs, the
parent system $I$ is isostatic (the root of the DR-plan), and we seek
\dfn{solution(s) (among a finite large number of solutions) with a
specific orientation or chirality}. In the case of underconstrained
qusecs the subsystems are the multiple roots of the DR-plan, the
parent system $I$ is underconstrained, and we typically seek an
efficient algorithmic description of \dfn{connected component(s) of
solutions with a specific orientation or chirality}.

The main barrier in recombination when given an optimal DR-plan (of
smallest possible size or max fan-in),  is that the number of children
(resp. number of roots) - and correspondingly the  size and complexity
of the (indecomposable) algebraic system $I$ to be solved - could be
arbitrarily large as a function of the size of the input system. This
difficulty can persist even after optimal parametrization of the
indecomposable system $I$ as in \uncited to minimize its algebraic
complexity.

In the following, we formulate the problem of \dfn{optimal
modification}, of the indecomposable algebraic system $I$ into a
decomposable system with a small DR-plan (low algebraic complexity).
Leveraging recent results on \dfn{Cayley configuration spaces}, our
approach to the optimal modification problem achieves the following:
\dfn{(a) Small DR-plan.} We obtain a  parameterized family of systems
$I_{\lambda_F}$ -  one for each value $\lambda_F$ for the parameters
$F$,  all of which have small DR-plans. Thus, given a value $v$ for
$\lambda_F$, the system $I_v$ can potentially be solved or realized
easily once the orientation type of the solution is known  (when the
DR-plan size is small enough). \dfn{(b) Solution preservation.}
Moreover, the union of solution spaces of the systems in the family
$I_{\lambda_F}$ is guaranteed to contain all of $I$'s solutions. \dfn{(c) Efficient search.} Finally, this union can be projected into the
so-called \dfn{Cayley} or distance parameter space $\lambda_F$ that is
convex or otherwise easy to traverse in order to search for $I$'s
solution (or connected component) of the desired orientation type. For
the case when the modification is bounded, this approach provides an
efficient algorithm for recombination. We first define the decision
version of the problem of optimal modification for decomposition. The
standard optimization versions are straightforward.

\problemstatement{Optimal Modification for Decomposition (OMD) Problem.}
Given a generically independent graph $G = (V,E)$ with no non-trivial
proper isostatic subgraph (indecomposable), and 2 constants $k$ and
$s$, does there exist a set of at most $k$  edges $E_1$ and a set of
nonedges $F$ such that the modified graph $H = (V, E\setminus E_1 \cup
F)$ has a DR-plan of size at most $s$?  The OMD$_k$ problem is  OMD
where $k$ is a fixed bound (not part of the input).   We say that such
a tuple $(G,s)$ is a member of the set OMD$_k$. We loosely refer to
graphs $G$ as OMD with appropriately small $k,s$ or OMD$_k$ graphs
with appropriately small $s$.

It is immediately clear that indecomposable graphs $G$ that belong in
OMD$_k$ for small $k$ and $s$  lend themselves to modification  into
decomposable graphs satisfying Criteria (a) and (b) above. However, it
is not clear how Criterion (c) is met by OMD graphs.
%
\subsection{Using Convex Cayley Configuration Spaces}
\label{sec:2-tree-reduction}
%
Next we provide the necessary background to describe a specific
approach for achieving the requirements (a)-(c) mentioned above, by
restricting the class of reduced graphs $G' = G\setminus E_1$ and
their isostatic completions $H$ in the above definition of the OMD
problem, and using a key theorem of Convex Cayley configuration spaces
\uncited. This theorem characterizes the class of graphs $H$ and non-
edges $F$ (Cayley parameters), such that the set of vectors
$\lambda_F$ of  attainable lengths of the nonedges $F$ is always
convex for any given lengths $\delta$ for the edges of $H$ (i.e, over
all the realization s of the bar-joint constraint system or linkage
$(H,\delta)$ in 2 dimensions). This set is called the (2-dimensional)
\dfn{Cayley configuration space} of the linkage $(H,\delta)$ on the
Cayley parameters $F$, denoted $\Phi_F(H,\delta)$ and can be viewed as
a ``projection'' of the cartesian realization space of $(H,\delta)$ on
the Cayley parameters $F$. Such graphs $H$ are said to have
\dfn{convexifiable Cayley configuration spaces with parameters $F$}
(short: \dfn{convexifiable}). To state the theorem, we first have to
define the notion of \dfn{2-sums} and \dfn{2-trees}. Let $H_1$ and
$H_2$ be two graphs on disjoint sets of vertices $V_1$ and $V_2$, with
edge sets $E_1$ and $E_2$ containing edges $(u,v)$ and $(w,x)$
respectively.  A \dfn{2-sum} of $H_1$ and $H_2$ is a graph $H$
obtained by taking the union of $H_1$ and $H_2$ and identifying $u=w$
and $v=w$. In this case, $H_1$ and $H_2$ are called \dfn{2-sum
components} of $H$. A \dfn{minimal 2-sum component} of $H$ is  one
that cannot be further split into 2-sum components. A \dfn{2-tree} is
recursively obtained by taking a 2-sum of two 2-trees, with the base
case of 2-tree being a triangle. A \dfn{partial 2-tree} is a 2-tree
minus some edges. Partial 2-trees are characterized as having $K_4$ as
forbidden minor are also called series parallel graphs.

\begin{theorem}\label{theorem:convexcayley}
    \uncited $H$ has a convexifiable Cayley configuration space  with
    parameters $F$ if and only if for each $f\in F$  all the minimal
    2-sum components of $H\cup F$ that contain both endpoints of $f$
    are partial 2-trees. The Cayley configuration space
    $\Phi_F(H,\delta)$ of a bar-joint system or linkage $(H,\delta)$
    is a convex polytope. When $H\cup F$ is a 2-tree, the bounding
    hyperplanes of this polytope are triangle inequalities relating
    the lengths of edges of the triangles in $H\cup F$.
\end{theorem}

@@

\begin{theorem}\label{theorem:omdk}
    Given an indecomposable graph $G$, let $G'$ be a spanning partial
    2-tree subgraph in $G$ with $k$ fewer edges than $G$. Then
    $(G,2)$ belongs in the set OMD$_k$.
\end{theorem}

@@

We refer to such graphs $G$ in short as \dfn{$k$-approximately
convexifiable}, where the reduced graphs $G'$ and isostatic
completions $H$ are convexifiable. As observed earlier, since graphs
such as $G$ are in OMD$_k$, Criteria (a) and (b) are automatically met
for small enough $k$. Criterion $c$ is addressed as described in the
following efficient search procedure which clarifies the dependence of
the complexity on the number and ranges  of Cayley parameters $F$.

\begin{theorem}\label{theorem:criterionc}
    [Efficient search]
    For an indecomposable, $k$-approximately convexifiable graph $G =
    (V,E)$, let $G' = (V,E' =E\setminus D)$ be a spanning partial
    2-tree subgraph where $|D| \le  k$. Let  $F$ be a set of nonedges
    of $G$ such that $H = (V, E'\cup F)$ is a 2-tree. Each solution
    $p$ (or connected component of a solution space) of $(G,\delta)$
    of an orientation type $\sigma_p$ can be found in time
    $O(\log(W))$ where $W$ is the number of cells of desired accuracy
    (discrete volume) of the convex polytope $\Phi_F(G',\delta_E')$.
    The (discrete) volume $W$ is exponential in $|F|$ and polynomial
    in the (discrete range) of the parameters in $F$.
\end{theorem}

\begin{proof}
    The Cartesian realization space of $(H,<\delta_E', \lambda_F>)$ is
    computed easily with a DR-plan of size 2, and is the union of
    $2^t$ solutions (modulo orientation preserving isometries) each
    with a distinct orientation type, where $t$ is the number of
    triangles in the 2-tree $H$; here $\delta_{S}$ is the restriction
    of the length vector $\delta$ to the edges in $S$. A desired
    solution $p$ (or connected component of a solution space) of
    $(G,\delta)$ of an orientation type $\sigma_p$ can be found by a
    subdivided binary search   of the cartesian realization space of
    $(H, <\delta_{E'},\lambda_F>)$ of orientation type $\sigma_p$ - as
    $\lambda_F$ ranges over the discretized convex polytope
    $\Phi_F(G',\delta_E')$ with bounding hyperplanes described in
    Theorem \ref{theorem:convexcayley}. A solution $p$  is found  when
    the lengths for nonedges in $D$ match $\delta_D$.
\end{proof}

@@

\noindent
\note A huge advantage of the convex Cayley method is that it is
completely unaffected when $\delta$ are intervals rather than exact
values~\cite{GaoSitharam}.
%cite my paper with Heping!!!

\medskip\noindent
\textbf{Example}. \todo{This uses different example format, not their environment. Should this, or section 3 change?}
A graph $G=K_{3,3}$  cannot be decomposed into any nontrivial
isostatic graphs, i.e, its DR-plan has a root and 9 children
corresponding to the 9 edges. Solving or recombining the system
$(G,\delta)$ corresponding to the root of this DR-plan involves
solving a quadratic system with 8 equations and variables. Instead of
simultaneously solving this system, we could instead use the fact that
$G=K_{3,3}$ is in OMD$_1$: remove the edge $e$ in Figure
\ref{fig:omd_k33_example} to give a partial 2-tree $G'$. Now add the
nonedge $f$ to give a 2-tree $H$ with a DR-plan of size 2. The Cayley
configuration space $\Phi_f(G', \delta_{E\setminus e})$ is a single
interval of attainable length values $\lambda_F$ for the edge $f$.
When $\delta$ is generic, i.e, does not admit collinearities or
coincidences in the realizations of $(G,\delta)$, the realization
space of $(H, <\delta_{E\setminus e}, \lambda_f>)$ has 16 solutions
$q^p_{\lambda_f}$ (modulo orientation preserving isometries), with
distinct orientation types $\sigma_p$ (two orientation choices for
each of the 4 triangles) that can be obtained by solving a sequence of
4 single quadratics in 1 variable (DR-plan of size 2). By subdivided
binary search in the interval $\lambda_f \in \Phi_f(G',
\delta_{E\setminus e})$, the desired solution $p$  of $(G,\delta)$ is
found when the length of the nonedge $e$ in  the realization
$q^p_{\lambda_f}$ is $\delta_e$.
%
\subsection{Optimized Modification by Enlarging the class of Reduced
Graphs}
\label{sec:tdecomp}
%
It is possible in principle to decrease $k$ for a OMD$_k$ graph (i.e,
the number of edges to be removed to ensure an isostatic completion
that is decomposable with a small DR-plan) by considering reduced
graphs $G'$ (and modified graphs $H$) that come from a larger class
than partial 2-trees but nevertheless have convex Cayley configuration
spaces at least when the realization space is restricted to a
sufficiently comprehensive orientation type. In particular, the so-
called \dfn{tree-decomposable graphs of low Cayley complexity}
\uncited include the partial 2-trees and many others that are not
partial 2-trees. These too result in DR-plans of size 2 or 3, putting
$G$ in the class OMD$_k$ and thus meeting Criteria (a) and (b). The
Criterion (c) is met---for example when $k=1$---because 1-dof Cayley
configuration spaces of linkages based on such graphs $G'$ are known
to be single intervals when a comprehensive orientation type
$\sigma_p$ of the sought solution $p$ is given. In addition, the
bounds of these intervals are of low algebraic complexity. More
precisely, the bounds  can themselves be computed using a DR-plan of
size 2 or 3, i.e, the computation of these bounds is tree-
decomposable. Alternatively, the bounds are in a simple quadratic or
radically solvable extension field over the rationals, or can be
computed by solving a triangularized system of quadratics.

### 2014-12-18 doc/GMP2015/s07_problempinning.tex
Updated the subdivision figure
Attempt to unify text decoration/emphasis. Make sure to pull, affects all tex files.

We can use DR-plans to help in solving questions in designs for graphene and silica bi-layers (\todo{Add Citation}). Specifically, we will look into the problem when you have a mono-layer and it is under-constrained. Our goal is to make such a layer isostatic. This can be done in a number of ways.

\begin{figure}\centering
    \includegraphics[width=0.4\linewidth]{img/Silica} \hspace{0.5cm}
    \raisebox{0.5\height}{\includegraphics[width=0.2\linewidth]{img/Silicon_tetrahedron}}
    \caption{Example of a single mono-layer of a Silica Oxygen glassy structure. This can be thought of as a multi-triangle-pin qusec where the Silicon atoms are the triangles and the Oxygen atoms are the pins. Not shown here are other layers stacked on this one. Each Silicon atom has the structure seen to the right, and so binds to another Oxygen atom in an adjacent mono-layer.}
    \label{fig:silica_glass}
\end{figure}

@@

    (b)
    \begin{subfigure}{0.6\linewidth}
        \includegraphics[width=\linewidth]{img/pentawesome}
    \end{subfigure}
    \caption{Examples of self-similarity in the form of repeated subdivision. In (a), a simple subdivision is done that does not guarantee isostatic. In (b), a more complicated scheme is used, ensuring that the resulting graph is both isostatic and non-tree-decomposable.}
    \label{fig:subdivision}
\end{figure}

@@

\begin{definition}
    A \emph{body-hyper-pin qusec} is a qusec where the object are rigid bodies that are pinned together by some number of pins.
\end{definition}

\begin{remark}
    A body-hyper-pin qusec is a special case of bar-joint qusecs that we have been studying. As such, the DR-planning discussed in Section \ref{sec:DRP} is unchanged and the work in Section \ref{sec:recomb} will still go through with slight modifications.
\end{remark}

@@

For the rest of this section, we will only be dealing with the DR-plan of such qusecs. As such, we will refer only to the underlying graph of the qusecs. We now go over 2 sub-classes of body-hyper-pin graphs for modeling Examples 4 and 5 in Section \ref{sec:intro}.

\begin{definition}
\label{def:body-pin}
    A body-pin graph is a body-hyper-pin graph with the following conditions:
    \begin{itemize}
        \item Each pin is shared by at most 2 bodies
        \item No 2 bodies share more than one pin
    \end{itemize}
    Such a body-pin graph can also be seen as a \emph{body-bar graph} where the bodies are the original bodies and each pin represents 2 bars from one body to another. Such body-bar graphs with 1 and 2-dof can be characterized by being $(3,4)$ and $(3,5)$-tight respectively.
\end{definition}

\begin{definition}
    A \emph{multi-triangle-pin graph} is a body-hyper-pin graph where each body is a triangle, meaning it is pinned in at most 3 places. This is also represented as a hyper-graph where each pin is a vertex and each triangle represents a tri-hyperedge. For such hyper-graphs, 1 and 2-dof can be characterized by $(2,4)$ and $(2,5)$ tightness respectively.
\end{definition}

@@

\begin{definition}
    A \emph{$(k,l)$-tight DR-plan} is one in which each child node is either a vertex maximal proper $(k,l)$-tight sub-graph of the parent node or it is trivial. In our case, the trivial nodes are just the bodies.
\end{definition}

@@

\begin{observation}
    For the 1-dof body-hyper-pin graphs described above, a DR-plan node will follow one of the following:

@@

\begin{theorem}
\label{thm:1dofcase}
    Given a 1-dof body-pin or multi-triangle graph $G$ and corresponding 1-dof DR plan $T$, there is a quadratic algorithm for the 1-dof optimal completion problem of Section \ref{sec:recomb} for body-pin and multi-triangle-pin graphs.
\end{theorem}

@@

XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX QUESTION XXXXXXXXXXXXXXXXXXXXXXXXX
An open question remains as to the complexity of the optimal completion problem when there are more than 2-dof. For higher number of dofs, the $(l,k)$ characterization is no longer matroidal.

>>>>

We can use DR-plans to help in solving questions in designs for graphene and silica bi-layers \uncited. Specifically, we will look into the problem when you have a mono-layer and it is under-constrained. Our goal is to make such a layer isostatic. This can be done in a number of ways.

\begin{figure}\centering
    \includegraphics[width=0.4\linewidth]{img/Silica} \hspace{0.5cm}
    \raisebox{0.5\height}{\includegraphics[width=0.2\linewidth]{img/Silicon_tetrahedron}}
    \caption{Example of a single mono-layer of a Silica Oxygen glassy structure. This can be thought of as a multi-triangle-pin qusecs where the Silicon atoms are the triangles and the Oxygen atoms are the pins. Not shown here are other layers stacked on this one. Each Silicon atom has the structure seen to the right, and so binds to another Oxygen atom in an adjacent mono-layer.}
    \label{fig:silica_glass}
\end{figure}

@@

    (b)
    \begin{subfigure}{0.8\linewidth}
        \includegraphics[width=\linewidth]{img/pentawesome}
    \end{subfigure}
    \caption{Examples of self-similarity in the form of repeated subdivision. In (a), a simple subdivision is done that does not guarantee isostatic. In (b), a more complicated scheme is used, ensuring that the resulting graph is both isostatic and non-tree-decomposable.}
    \label{fig:subdivision}
\end{figure}

@@

\begin{definition}
    A \dfn{body-hyperpin qusecs} is a qusecs where the object are rigid bodies that are pinned together by some number of pins.
\end{definition}

\begin{remark}
    A body-hyperpin qusecs is a special case of bar-joint qusecs that we have been studying. As such, the DR-planning discussed in Section \ref{sec:DRP} is unchanged and the work in Section \ref{sec:recomb} will still go through with slight modifications.
\end{remark}

@@

For the rest of this section, we will only be dealing with the DR-plan  of such qusecs. As such, we will refer only to the underlying graph of the qusecs. We now go over 2 sub-classes of body-hyperpin graphs for modeling Examples 4 and 5 in Section \ref{sec:intro}.

\begin{definition}
\label{def:body-pin}
    A body-pin graph is a body-hyperpin graph with the following conditions:
    \begin{itemize}
        \item Each pin is shared by at most 2 bodies
        \item No 2 bodies share more than one pin
    \end{itemize}
    Such a body-pin graph can also be seen as a \dfn{body-bar graph} where the bodies are the original bodies and each pin represents 2 bars from one body to another. Such body-bar graphs with 1 and 2-dof can be characterized by being $(3,4)$ and $(3,5)$-tight respectively.
\end{definition}

\begin{definition}
    A \dfn{multi-triangle-pin graph} is a body-hyperpin graph where each body is a triangle, meaning it is pinned in at most 3 places. This is also represented as a hyper-graph where each pin is a vertex and each triangle represents a tri-hyperedge. For such hyper-graphs, 1 and 2-dof can be characterized by $(2,4)$ and $(2,5)$ tightness respectively.
\end{definition}

@@

\begin{definition}
    A \dfn{$(k,l)$-tight DR-plan} is one in which each child node is either a vertex maximal proper $(k,l)$-tight sub-graph of the parent node or it is trivial. In our case, the trivial nodes are just the bodies.
\end{definition}

@@

\begin{observation}
    For the 1-dof body-hyperpin graphs described above, a DR-plan node will follow one of the following:

@@

\begin{theorem}
\label{thm:1dofcase}
    Given a 1-dof body-pin or multi-triangle graph $G$ and corresponding 1-dof DR-plan $T$, there is a quadratic algorithm for the 1-dof optimal completion problem of Section \ref{sec:recomb} for body-pin and multi-triangle-pin graphs.
\end{theorem}

@@

\begin{openproblem}
    What is the complexity of the optimal completion problem when there are more than 2-dof. For higher number of dofs, the $(k,l)$ characterization is no longer matroidal.
\end{openproblem}
